A Partitioning Algorithm for Maximum Common Subgraph Problems\thanks{This work was supported by a pile of money from a funding agency.}

### 2017-02-16 paper/paper.tex
Use booktabs
Thinging

\usepackage{ijcai17}
\usepackage{times}
\usepackage{graphicx}
\usepackage{cleveref}
\usepackage{microtype}

@@

      #2
  \end{minipage}
  \quad
  \begin{minipage}[t]{0.3\linewidth}
      \centering Labelling of $G$
      \begin{tabular}[t]{cc}
      \hline
          Vertex & Label\\
      \hline
          #3
      \hline
      \end{tabular}
  \end{minipage}
  \quad
  \begin{minipage}[t]{0.3\linewidth}
      \centering Labelling of $H$
      \begin{tabular}[t]{cc}
      \hline
          Vertex & Label\\
      \hline
          #4
      \hline
      \end{tabular}
      \medskip
  \end{minipage}
}

@@

The
\McSplit\ algorithm may be considered as a version of the constraint
programming, forward-checking (CP-FC) algorithm of
\citet{DBLP:conf/cp/NdiayeS11}, using an efficient, compressed representation
of domains.  In the CP-FC algorithm, each vertex in $v \in \V(G)$ is
represented by a variable, whose domain corresponds to the set of vertices in
$\V(H)$ to which $v$ may currently be mapped, with an additional special $\bot$
value representing an unmapped vertex.

Given a label class $\langle P,T \rangle$ in \McSplit\, the vertices in $P$
correspond to variables in CP-FC, all of which have domain equal to $T \cup \{
    \bot \}$.  The label-class representation of domains is possible because
throughout the CP-FC algorithm for max common subgraph, the domains of any two
variables are either identical or disjoint.  (To the best of our knowledge,
this observation has not been made previously.)

Where CP-FC uses a soft allDifferent constraint to filter and compute a bound,
\McSplit\ computes a bound from each label class in the future, i.e.\ it
performs Hall Set reasoning on the fly.

\McSplit\ requires $\BigO{g+h}$ time at each search node, and does not require
any auxiliary data structures larger than the adjacency matrices of $G$ and $H$
to be built prior to search.  CP-FC requires $\BigO(?????)$ \dots The clique
encoding requires an association graph with $g h$ vertices to be
constructed before search.  Modern maximum-clique algorithms typically perform
a colouring step at each search node that is quadratic in the order of the
graph, and thus $\BigO{g^2h^2}$ in the order of the max common subgraph input
graphs.

?? Mention partition backtracking

?? Mention dual viewpoint heuristics

The correctness of \McSplit\ can be seen by noting its equivalence to CP-FC
\citep{DBLP:conf/cp/NdiayeS11}, if we replace the label-class representation of
$\mathit{future}$ with domain stores in \cref{McCliqueAlg}.  To obtain
the CP-FC algorithm, replace \linerangeref{InnerLoop}{InnerLoopEnd} of
\cref{McCliqueAlg} with \cref{cpAlg}.  The pruning of domains
carried out in these lines can be easily seen to be equivalent to the
corresponding lines in \cref{McCliqueAlg}.

\begin{algorithm}
\DontPrintSemicolon
\nl    \For {$u \in \N(G, b)$}{
\nl        remove $v$ from $D'_u$ \;
\nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
       }
\nl    \For {$u \in \invN(G, b)$}{
\nl        remove $v$ from $D'_u$ \;
\nl        remove the neighbourhood of $t$ from $D'_u$ \;
       }
\caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm}
\label{cpAlg}
\end{algorithm}

TO WRITE PROPERLY: The second difference is in the bound function: the French
code constructs a maximum cardinality matching, with some special weighting
trickery for $\bot$, whilst James just sums the smaller of the two sides of
each partition. If you think a bit about what the matching looks like, you can
see that these two give the same answer, but at very different speeds.

\paragraph{CP-FC and \McSplit\ search nodes on MCS instances}
\cref{figure:33ved-james-versus-cp-fc-nodes-scatter} shows one point per instance,
with search nodes visited by CP-FC on the horizontal axis and search nodes visited
by \McSplit\ on the vertical axis.  There is a strong correlation, which reflects
the similarity of the search space explored by the two algorithms.  \McSplit\
tends to explore slightly fewer search nodes than CP-FC, because the label-classes
allow us to (cheaply) compute a very good heuristics for variable selection.

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
    \caption{MCS 33\% vertex and edge labelled directed instances, James vs
    CP-FC, Nodes (TODO?? heatmapify this and exclude
    timeouts)}\label{figure:33ved-james-versus-cp-fc-nodes-scatter}
\end{figure}

\paragraph{kdown and \McSplit\ search nodes on SIP instances}
\cref{figure:sip-james-versus-kdown-nodes-scatter} shows the number of search
nodes taken by kdown and \McSplit\ on each instance.  Although \McSplit\ is the
faster algorithm overall, it explores more search nodes than kdown for most instances.
The scatter plot has a concentration of points around the line $y=2x$. We conjecture
that the two algorithms explore very similar search trees for instances close to this
line.

>>>>

\usepackage{ijcai17}
\usepackage{times}
\usepackage{graphicx}
\usepackage{cleveref}
\usepackage{microtype}
\usepackage{booktabs}

@@

      #2
  \end{minipage}
  \quad
  \begin{minipage}[t]{0.3\linewidth}
      \centering
      Labelling of $G$
      \begin{tabular}[t]{cc}
      \toprule
          Vertex & Label\\
      \midrule
          #3
      \bottomrule
      \end{tabular}
  \end{minipage}
  \quad
  \begin{minipage}[t]{0.3\linewidth}
      \centering
      Labelling of $H$
      \begin{tabular}[t]{cc}
      \toprule
          Vertex & Label\\
      \midrule
          #4
      \bottomrule
      \end{tabular}
      \medskip
  \end{minipage}
}

@@

Our results so far suggest that \McSplit\ has broadly similar performance
trends to the constraint programming, forward-checking (CP-FC) algorithm of
\citet{DBLP:conf/cp/NdiayeS11}, but with much lower constant factors. Indeed,
in \cref{figure:33ved-james-versus-cp-fc-nodes-scatter} we plot the number of
recursive calls made by our algorithm on the $y$-axis versus the number made by
CP-FC on the $x$-axis, for the labelled, unconnected problem instances. We see
a close correlation: we typically do slightly less work, and sometimes do more,
but instances with more than one order of magnitude difference in search space
size are rare.

Why is this? We do not see a similar correlation between our search space size
and that of the clique spproach. The key observation is that \McSplit\ may be
considered to be a different version of the constraint CP-FC algorithm, using
an unconventional domain store and more efficient filtering algorithms. We now
explore this relationship further.

\begin{figure}[t]
    \centering
    \includegraphics*{gen-graph-33ved-james-versus-cp-fc-nodes-scatter.pdf}
    \caption{Size of the search space for \McSplit versus CP-FC on the
    labelled, directed, unconnected instances which were solved within the
    timeout by both algorithms.}
    \label{figure:33ved-james-versus-cp-fc-nodes-scatter}
\end{figure}

In the CP-FC algorithm, each vertex in $v \in \V(G)$ is represented by a
variable, whose domain corresponds to the set of vertices in $\V(H)$ to which
$v$ may currently be mapped, with an additional special $\bot$ value
representing an unmapped vertex.
Given a label class $\langle P,T \rangle$ in \McSplit\, the vertices in $P$
correspond to variables in CP-FC. The label-class representation of domains is
possible because throughout the CP-FC algorithm for max common subgraph, the
domains of any two variables are either identical or disjoint (excluding
$\bot$, which is slightly more complicated). To the best of our knowledge, this
observation has not been made previously, and it is not exploited in other
implementations.

CP-FC uses a soft all-different constraint to compute a bound, which requires
running a matching algorithm on a supporting compatibility graph.  Careful
thought shows that \McSplit\ computes the same bound, but using a simple
counting loop---this is only possible because of the disjoint nature of the
domains.

%%% This equivalence to CP-FC makes the correctness of \McSplit\ easy to establish:
%%% if we replace the label-class representation of $\mathit{future}$ with
%%% conventional domain stores, and replace \linerangeref{InnerLoop}{InnerLoopEnd}
%%% of \cref{McCliqueAlg} with \cref{cpAlg}.  The pruning of domains carried out in
%%% these lines can be easily seen to be equivalent to the corresponding lines in
%%% \cref{McCliqueAlg}.
%%% 
%%% 
%%% \begin{algorithm}
%%% \DontPrintSemicolon
%%% \nl    \For {$u \in \N(G, b)$}{
%%% \nl        remove $v$ from $D'_u$ \;
%%% \nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
%%%        }
%%% \nl    \For {$u \in \invN(G, b)$}{
%%% \nl        remove $v$ from $D'_u$ \;
%%% \nl        remove the neighbourhood of $t$ from $D'_u$ \;
%%%        }
%%% \caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm ?? Not sure about this bit, maybe we don't need it}
%%% \label{cpAlg}
%%% \end{algorithm}

A further advantage of our encoding is that it gives us efficient access to a
better branching heuristic. The CP-FC algorithm uses smallest domain first,
which in our algorithm corresponds to branching on a label class with smallest
$|T|$. We instead branch on the label class with smallest $\min(|P|,|T|)$.
This is empirically better, and accounts for much of the difference between
the number of recursive calls made; branching on the smallest $|P| |T|$ gives
very similar results. This can be viewed as exploiting both smallest domain first,
and the dual viewpoint \citep{DBLP:conf/ecai/Geelen92} of smallest domain
first, simultaneously, but we do not have the overheads of having to maintain
and channel between the dual viewpoint that would be required when using a
conventional domain store. ?? Check this!

What about our relationship to the $k\downarrow$ of \citet{UpcomingAAAIPaper}?
\Cref{figure:sip-james-versus-kdown-nodes-scatter} plots the number of
recursive calls made by $k\downarrow$ and \McSplit\ on each of the subgraph
isomorphism instances. Although \McSplit\ is the faster algorithm overall, it
explores more search nodes than $k\downarrow$ for most instances.

?? Talk about the classic tradeoff, and dynamically switching.

?? The concentration of points, down to unit propagation and counting
